The dependence of spiral disk morphology on the star formation-stellar mass relation

### 2014-08-11 mainsequence_morph.tex
Updated fonts and captions for figures

\begin{abstract}
We measure the mass-star formation relation in disk galaxies at $z<0.05$, using Galaxy Zoo morphologies to examine different populations of spirals as classified by their kiloparsec-scale structure. We examine the number of arms, their average pitch angle, and the presence of a galactic bar, and show that both the slope and dispersion of the SFR-$M_\star$ relation is constant when varying all the above parameters. We interpret this as evidence that the spiral arms, which are imperfect reflections of the galaxy's current gravitational potential, are either fully independent of the basic quenching relation or are completely overwhelmed by the combination of outflows and feedback. The arrangement of the star formation can be changed (as demonstrated by the filling factor of the disk), but the system \emph{as a whole} regulates itself. We also show that mergers (both major and minor), which represent the strongest conditions for increases in star formation at a constant mass, only drive the galaxies off the main relation by $\sim0.3$~dex; this is a significant reduction over merging systems at $z>1$. 
\end{abstract}

@@

As star-forming galaxies exhibit a wide range of physical appearances in optical images, we can ask the natural question of whether the specifics of this physical appearance, and its underlying dynamical processes, have any effect on this homeostasis and therefore the galaxy's location relative to the SFMS. If the details of a galaxy's physical appearance are correlated with position relative to the main sequence, then the dynamical processes that give rise to them -- such as bar formation and spiral arm pitch angle -- are a fundamental aspect of the galaxy's regulatory mechanism. If, on the other hand, these features are not correlated, then there are two options: either galaxy substructure is simply not relevant to the overall $M_\star$-SFR relationship, or the regulatory mechanism overcomes the local effect of substructure in all circumstances. This would speak to the strength of the regulator.

@@

%How many galaxies are there in the total sample? What is the mass, color, and redshift range?
The total sample analyzed contains 52,685 star-forming galaxies. These are selected from the GZ2 spectroscopic sample with $z<0.1$, $M_r<-19.5$, and classified as actively star-forming (BPT$=1$) from the MPA-JHU emission line measurements. The average color for the star-forming galaxies is relatively blue, with $(u-r)=1.7\pm0.6$.  Subsets of the star-forming galaxies are listed in Table~\ref{tbl-fits}.

@@

\emph{Come up with a more sensible measure of the uncertainty - the formal covariance elements clearly don't represent the dispersion in the main sequence. Interquartile range?}

@@

\begin{figure*}
\includegraphics[angle=0,width=7.0in]{figures/ms_arms_number.pdf}
\caption{Stellar mass vs. star formation rate; grayscale colors are the distribution of all star-forming galaxies from the MPA-JHU catalog. Colored points in each panel show spiral galaxies with ``tight'', ``medium'', and ``loose'' spiral arms as identified by morphology flags in GZ2. Dotted lines show the weighted least-squares linear fit to the population as split by pitch angle; the solid line is the fit to all star-forming galaxies. 
\label{fig-number}}
\end{figure*}

@@

\begin{figure*}
\includegraphics[angle=0,width=7.0in]{figures/ms_arms_winding.pdf}
\caption{Stellar mass vs. star formation rate; grayscale colors are the distribution of all star-forming galaxies as in Figure~\ref{fig-number}. From left to right: red, green, and blue points spiral galaxies with ``tight'', ``medium'', and ``loose'' spiral arms as identified by morphology flags in GZ2. Dotted lines show the weighted least-squares linear fit to the population as split by pitch angle; the solid line is the fit to all star-forming galaxies. 
\label{fig-winding}}
\end{figure*}

@@

\begin{figure*}
%\includegraphics[angle=0,width=7.0in]{figures/ms_bar.pdf}
\includegraphics[angle=0,width=7.0in]{figures/ms_bar_contour.pdf}
\caption{Stellar mass vs. star formation rate; grayscale colors are the distribution of all star-forming galaxies as in Figure~\ref{fig-number}. Left: blue contours show the distribution of barred galaxies ($p_\textrm{bar}\ge0.4$ for previously-identified disks) from GZ2. Right: red contours are the distribution of remaining disk galaxy population with no evidence for a strong bar ($p_\textrm{bar}<0.4$). Dotted lines show the weighted least-squares linear fit to the barred/unbarred population; the solid line is the fit to all star-forming galaxies. 
\label{fig-bar}}
\end{figure*}

@@

\begin{figure*}
\includegraphics[angle=0,width=7.0in]{figures/ms_mergers_both.pdf}
\caption{\textbf{Left:} Stellar mass vs. star formation rate as in Figure~\ref{fig-number} for 2,978~merging galaxies in GZ, compared to the broader star-forming population. Merging systems are colour-coded by the mass ratio of the primary-to-secondary galaxy; there is no clear difference in the merging populations with regard to the SFMS when comparing major to minor mergers. When fixing the slope of the star-forming main sequence and allowing the offset to vary, mergers have higher SFRs by $\sim0.3$~dex. \textbf{Right:} Same as left plot, but binned and colour-coded by merger fraction ($N_\textrm{mergers}/N_\textrm{star-forming galaxies})$. Overplotted linear fits are same as before. This demonstrates that the galaxies that lie furthest above the SFMS are $>50\%$ mergers. 
\label{fig-mergers}}
\end{figure*}

@@

The most significant forcing event to the galaxy system known -- a major galaxy merger -- has only a slight effect. Star formation rates are increased by an average of 0.29~dex (less than a factor of 2) in major mergers (Figure~\ref{fig-mergers}). \citet{dar10} showed that in the local ($z<0.1$) Universe, galaxies involved in major mergers do induce intense bursts of star formation, but specifically in only the spiral (disc) galaxies taking part and increasing the local sSFR by only a factor $\sim2$. This increase in star formation is much stronger in galaxies at $1.5<z<2.5$, likely due to the higher gas fractions \citep{dad10,rod11}.

@@

\clearpage

>>>>

\begin{abstract}
We measure the mass-star formation relation in disk galaxies at $z<0.05$, using Galaxy Zoo morphologies to examine different populations of spirals as classified by their kiloparsec-scale structure. We examine the number of spiral arms, their average pitch angle, and the presence of a galactic bar in the disk, and show that both the slope and dispersion of the $M_\star$-SFR relation is constant when varying all the above parameters. We also show that mergers (both major and minor), which represent the strongest conditions for increases in star formation at a constant mass, only drive the galaxies off the main relation by $\sim0.3$~dex; this is a significant reduction over the increase seen in merging systems at $z>1$. Of the galaxies lying significantly above the $M_\star$-SFR relation in the local Universe, more than $50\%$ are mergers. We interpret this as evidence that the spiral arms, which are imperfect reflections of the galaxy's current gravitational potential, are either fully independent of the basic quenching relation or are completely overwhelmed by the combination of outflows and feedback. The arrangement of the star formation can be changed (as demonstrated by the filling factor of the disk), but the system \emph{as a whole} regulates itself even in the presence of strong dynamical forcing. 
\end{abstract}

@@

As star-forming galaxies exhibit a wide range of physical appearances in optical images, we can ask the natural question of whether the specifics of this physical appearance, and its underlying dynamical processes, have any effect on this homeostasis and therefore the galaxy's location relative to the SFMS. If the details of a galaxy's physical appearance are correlated with position relative to the main sequence, then the dynamical processes that give rise to them -- such as bar formation and spiral arm pitch angle -- are a fundamental aspect of the galaxy's regulatory mechanism. If, on the other hand, these features are not correlated, then there are two options: either galaxy substructure is simply not relevant to the overall $M_\star$-SFR relationship, or the regulatory mechanism overcomes the local effect of substructure in all circumstances. This ultimately relates to the physical processes that control the overall strength of the regulator in each galaxy.

@@

%How many galaxies are there in the total sample? What is the mass, color, and redshift range?
The total sample analyzed contains 52,685 star-forming galaxies. These are selected from the GZ2 spectroscopic sample with $z<0.1$, $M_r<-19.5$, and classified as actively star-forming ($BPT=1$) from the MPA-JHU emission line measurements. The average color for the star-forming galaxies is relatively blue, with $(u-r)=1.7\pm0.6$.

@@

@@

\begin{figure*}
\includegraphics[angle=0,width=7.0in]{figures/ms_arms_number.pdf}
\caption{Total star formation rate as a function of stellar mass; grayscale colors are the distribution of all star-forming galaxies in SDSS from the MPA-JHU DR7 catalog. Colored points in each panel show spiral galaxies with 1, 2, 3, 4, more than four, or ``uncertain'' numbers of spiral arms per galaxy, as identified by GZ2 morphology flags. Dotted lines show the weighted least-squares linear fit to each population as split by arm multiplicity; the solid line is the fit to all of the star-forming galaxies. 
\label{fig-number}}
\end{figure*}

@@

\begin{figure*}
\includegraphics[angle=0,width=7.0in]{figures/ms_arms_winding.pdf}
\caption{Total star formation rate as a function of stellar mass; grayscale colors are the same as in Figure~\ref{fig-number}. From left to right: red, green, and blue points are spiral galaxies with ``tight'', ``medium'', and ``loose'' winding spiral arms as identified by GZ2 morphology flags. Dotted lines show the weighted least-squares linear fit as split by pitch angle; the solid line is the fit to all star-forming galaxies. 
\label{fig-winding}}
\end{figure*}

@@

\begin{figure*}
%\includegraphics[angle=0,width=7.0in]{figures/ms_bar.pdf}
\includegraphics[angle=0,width=7.0in]{figures/ms_bar_contour.pdf}
\caption{Total star formation rate as a function of stellar mass; grayscale colors are the same as in Figure~\ref{fig-number}. Left: blue contours show the distribution of barred galaxies ($p_\textrm{bar}\ge0.4$ for previously-identified disks) from GZ2. Right: red contours are the distribution of remaining disk galaxy population with no evidence for a strong bar ($p_\textrm{bar}<0.4$). Dotted lines show the weighted least-squares linear fit to the barred/unbarred population; the solid line is the fit to all star-forming galaxies. 
\label{fig-bar}}
\end{figure*}

@@

\begin{figure*}
\includegraphics[angle=0,width=7.0in]{figures/ms_mergers_both.pdf}
\caption{Total star formation rate as a function of stellar mass; grayscale colors are the same as in Figure~\ref{fig-number}. Left: Coloured points show 2,978~merging galaxies from \citet{dar10a}. Mergers are colour-coded by the mass ratio of the primary and secondary galaxies; there is no clear difference in the merging populations with regard to the SFMS when comparing major to minor mergers. When fixing the slope of the star-forming main sequence and allowing the offset to vary, mergers (dotted line) have higher SFRs by $\sim0.3$~dex compared to all star-forming galaxies (solid line). Right: Star-forming galaxies binned and colour-coded by merger fraction ($N_\textrm{mergers}/N_\textrm{star-forming galaxies})$. Overplotted lines are the same as left plot. Of the galaxies that lie furthest above the SFMS, more than $50\%$ are unambiguous mergers. 
\label{fig-mergers}}
\end{figure*}

@@

As a comparison to the kpc-scale structures discussed above, we analyze the impact of the most significant forcing event to a galaxy system known -- a major galaxy merger (Figure~\ref{fig-mergers}). In these systems (which are in various stages of coalescence) the star formation rates are increased by only an average of 0.29~dex, corresponding to less than a factor of two overall. \citet{dar10} showed that in the local ($z<0.1$) Universe, galaxies with intense bursts of star formation are limited to only the spiral (disc) galaxies. This increase in star formation for mergers does show a strong evolution in redshift out to at least $1.5<z<2.5$, likely due to the higher gas fractions involved \citep{dad10,rod11}.

@@
